Aggressive Scheduling for Numerical Programs

### 2012-04-30 report/draft.tex
Add software pipelining introduction.

Our approach gives us several advantages. The first one is that we are able to generate fully optimized
code using software pipelining from high level description. The second and final advantage, is that
we are still able to use the nice features provided by the C compiler (specially portability)
while tuning the scheduling parameters.

@@

\subsection{Related Work}
\begin{verbatim}
+( ) OSKI
+( ) FFTW
+( ) ATLAS/PhiPack
+( ) [Yotov] discusses how ATLAS only needed fine tuning and many of the course
  adjustments could be determined analytically.
+( ) Lam's software pipelining technique
\end{verbatim}

@@

\begin{verbatim}
+( ) A general technique(hack) for forcing schedules in C
+( ) An implemented set of rules in SPIRAL for tunable software pipelining.
+( ) And as a result, potentially faster code that is generated by SPIRAL that is
  comparable to an expert.
\end{verbatim}

@@

\subsection{Software Pipelining}
\begin{verbatim}
+( ) Go into the background of software pipelining. How is it done?
+( ) How do we do it in SPIRAL?
     -( ) What parameters can we tune? (how much to delay an op)
     -( ) Formally what does this look like? What is our algorithm?
\end{verbatim}

>>>>

Our approach gives us several advantages. The first one is that we are able to generate fully optimized
code using software pipelining from a high level description. The second and final advantage, is that
we are still able to use the nice features provided by the C compiler (specially portability)
while tuning the scheduling parameters.

@@

\subsection{Related Work}

\begin{verbatim}
+( ) OSKI
+( ) FFTW
+( ) ATLAS/PhiPack
+( ) [Yotov] discusses how ATLAS only needed fine tuning and many of the course
  adjustments could be determined analytically.
+( ) Lam's software pipelining technique
\end{verbatim}

Lam's software pipelining technique~\cite{Lam:04}.

@@

\begin{comment}
+( ) A general technique(hack) for forcing schedules in C
+( ) An implemented set of rules in SPIRAL for tunable software pipelining.
+( ) And as a result, potentially faster code that is generated by SPIRAL that is
  comparable to an expert.
\end{comment}

@@

\subsection{Software Pipelining}

Software pipelining is a coding technique that optimizes loops by overlapping instructions
from different loop iterations in order to exploit instruction level parallelism.
Due to this overlap we say that software pipelining is a type of out-of-order execution.
While software pipelining is somewhat similar to hardware pipelining, the order of the executions
is not determined by the processor but by the compiler or the programmer.

With software pipelining, we can reduce
pipeline stalls in the hardware, since the pipeline can be fed with new instructions from
different loop iterations. This greatly increases instruction parallelism because we can
start a new loop iteration before the past ones have completed. For this technique to be
effective, each loop is transformed into three sections:

\begin{description}
   \item[Prologue:] We start adding several iterations to the hardware pipeline.
   This phase ends when the last instruction of the first instruction is to be executed.

   \item[Steady State:] This phase starts where the prologue ends. Most execution time is spent
   here and this finishes when the first instruction of the last iteration is executed.

   \item[Epilogue:] In this phase, the hardware pipeline executes the last remaining instructions.
\end{description}

The need to duplicate the loop code into three phases is a notorious disadvantage of software
pipelining, because it may increase the program's code size. Moreover, since software pipelining
is usually used in conjunction with loop unrolling, this enlarges the code even more.

\begin{verbatim}
+( ) Go into the background of software pipelining. How is it done?
+( ) How do we do it in SPIRAL?
     -( ) What parameters can we tune? (how much to delay an op)
     -( ) Formally what does this look like? What is our algorithm?
\end{verbatim}
